### 2022-06-20 manuscript/manuscript.Rmd
advances with manuscript

Data is presented as median (*Mdn*) and interquartile range (*IQR*) for continuous variables, and as absolute and relative sample size for categorical/discrete variables.

A non-parametric approach was used since the underlying distribution of continuous outcomes, evaluated through analytical and graphical methods, did not follow a Gaussian distribution.

@@

Generalized additive models (GAM) were used to describe linear and non-linear relationships between developmental characteristics through smooth terms that are represented using penalized regression splines [@wood2011fast]. The restricted maximum likelihood estimation method was used for the selection of the smoothing parameter, and the thin-plate regression spline as the smoothing basis method, as they are the optimal smoother of any given basis dimension/rank [@wood2003thin]. GAMs were fitted using the mcgv package [@wood2017generalized]. To describe the smooth term, we use approximative derivative to summarise de trend in terms of linear segments.

[...], using the *R* programming language for statistical computing, [@rlanguage] and complementary R packages for model fitting, visualization and complementary computations [@wood2017generalized; @hadley2016ggplot2; @daniel2021see; @dominique2020estimation].

@@

# Overall slope
gam_coef <- report_slope(mod, trend = c("edad_cronologica_meses"))

# Linear segments
modelbased::estimate_smooth(
  data = modelbased::estimate_relation(mod, length = 100),
  x = "edad_cronologica_meses"
)

@@

When modelling the effect of chronological age on developmental skills, we found an averaged negative marginal effect on the communication score (`r gam_coef`), whereas from 1 to 5.7 moths old

>>>>

Data is presented as median (*Mdn*) and interquartile range (*IQR*) for continuous variables; for categorical/discrete variables, the absolute and relative sample size was reported.

A non-parametric approach was used since the underlying distribution of continuous measured outcomes, assessed through analytical and graphical methods, did not follow a Gaussian distribution.

@@

Generalized additive models (GAM) were used to describe linear and non-linear relationships in the form of smooth terms between developmental characteristics, represented through penalized regression splines [@wood2011fast]. Restricted maximum likelihood (i.e., REML) method was used for the estimation of the smoothing parameters, and thin-plate regression splines for smoothing basis, as they are the optimal smoother of any given basis dimension/rank [@wood2003thin]. To describe the smooth terms by means of quasi-linear segments, we used approximative effect derivatives to summarise the trend with 95% confidence intervals.

A probability of committing a type I ($\alpha$) error of less than 5% (*p* < 0.05), was considered sufficient evidence for statistical significance in hypothesis testing. The *R* programming language was used for statistical computing [@rlanguage]. GAMs and the corresponding model estimates were computed using the *R* packages *mcgv* and *modelbased* [@wood2017generalized; @dominique2020estimation]. Complementary R packages were used for visualization purposes [@hadley2016ggplot2; @daniel2021see].

@@

# Overall slope
(gam_coeffs <- report_overall(mod, "edad_cronologica_meses"))
(gam_slopes <- report_slopes(mod, "edad_cronologica_meses"))
(gam_smooth <- report_smooth(mod))

@@

When modelling the effect of chronological age on developmental skills, we found a significant non-linear effect (`r gam_smooth`), that reflect an averaged negative marginal effect on communication skills (`r gam_coeffs`), however, this wasn't true when assessing the direction of the effect in the age range between 1 to 7.6 (`r gam_slopes[1, expr]`), and neither in the 18 to 48 months old group (`r gam_slopes[3, expr]`), whereas the effect tend to be positive but non-significant.

```{r eval=FALSE}
mod <- gam(comunicacion_total ~ s(edad_corregida_meses), data = dataset, method = "REML")

# Overall slope
(gam_coeffs <- report_overall(mod, "edad_corregida_meses"))
(gam_slopes <- report_slopes(mod, "edad_corregida_meses"))
(gam_smooth <- report_smooth(mod))

plot_gam(dataset, edad_corregida_meses, comunicacion_total, round_to = 5,
         "Corrected age (months)", "Communication score", method = "REML")
```
